---
title: "Code"
output:
  html_document:
    number_section: true
    pandoc_args: [--number-offset=9]
---


```{r For setup, echo = FALSE, include = FALSE}
# Working directory ... set your favorite
setwd("C:/cygwin/home/NOC/ebook/edoc")

# Character width in console output
options(width = 84)

# Plot related setting
SAVE_PLOT_PDF <- F
if (SAVE_PLOT_PDF == TRUE){
  # Exporting to PDF file
  pdf(height = 7 / (4/3))

  # Rasterize the font information
  require(showtext)
  font.add("meiryo", "meiryo.ttc")
  par(family = "meiryo")
  showtext.begin()
}
```




#Batch solution for general state space model

##MCMC

```{r Figure 10.1, echo = FALSE, results='hide'}
# <<Convergence example of Markov chain>>
 
# Refer to https://rstudio-pubs-static.s3.amazonaws.com/29244_555e5b85c5ba49f384d4ab502dd96fae.html

# Preprocessing
require(rstan)
set.seed(1234)

# Non-convergence example
non_converged_model <- "
  parameters{
    real a;
    real b;
  }

  model{
    0 ~ normal(a + b, 1);
  }
"
stan_fit_tmp <- stan(model_code = non_converged_model, seed = 12343)
stan_out_tmp <- rstan::extract(stan_fit_tmp, permuted = FALSE)[, , "a"]

sample_min <- min(stan_out_tmp)
sample_max <- max(stan_out_tmp)
sample_len <- dim(stan_out_tmp)[1]
sample_var <- dim(stan_out_tmp)[2]
dens_len <- 512

x_val <- density(stan_out_tmp[, 1], from = sample_min, to = sample_max)$x
y_val <- sapply(1:sample_var, function(i){
           density(stan_out_tmp[, i], from = sample_min, to = sample_max)$y
         })

split.screen(rbind(c(0, 0.7, 0, 1), c(0.7, 1, 0, 1)))
chain <- c("dotted", "dashed", "blank", "solid")

screen(1); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 2, 2, 1))
for (i in c(1, 2, 4)){
  if (i != 1){ par(new = TRUE) }
  plot(stan_out_tmp[, i], type = "l", lty = chain[i], ann = FALSE, axes = FALSE,
       xlim = c(1, sample_len), ylim = c(sample_min, sample_max))
}
axis(side = 1, xlim = c(1, sample_len))
mtext(text = "Number of searching steps", side = 1, line = 2)
mtext(text = "Sample value", side = 2, line = 0)
legend("topright", legend = c("1st", "2nd", "3rd"), lty = chain[c(1, 2, 4)], cex = 0.6)

screen(2); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 0, 2, 0))
plot(0, type = "n", ann = FALSE, axes = FALSE,
     xlim = c(min(y_val), max(y_val)*2.0), ylim = c(sample_min, sample_max*1.1))
for (i in c(1, 2, 4)){
  polygon(y = c(x_val, rev(x_val)), x = c(y_val[, i], rep(y_val[dens_len, i], dens_len)),
          border = NA, col = rgb(0.5, 0.5, 0.5, 0.5))
  lines(x = y_val[, i], y = x_val, lty = chain[i])
}
text(x = max(y_val), y = median(x_val), adj = -0.2, labels = "Distribution of interest", srt = -90)

close.screen(all = TRUE)

# Convergence example
converged_model <- "
  parameters{
    real<lower=0> a;
    real<lower=0> b;
  }

  model{
    0 ~ normal(a + b, 1);
  }
"
stan_fit_tmp <- stan(model_code = converged_model, seed = 12343)
stan_out_tmp <- rstan::extract(stan_fit_tmp, permuted = FALSE)[, , "a"]

sample_min <- min(stan_out_tmp)
sample_max <- max(stan_out_tmp)
sample_len <- dim(stan_out_tmp)[1]
sample_var <- dim(stan_out_tmp)[2]
dens_len <- 512

x_val <- density(stan_out_tmp[, 1], from = sample_min, to = sample_max)$x
y_val <- sapply(1:sample_var, function(i){
           density(stan_out_tmp[, i], from = sample_min, to = sample_max)$y
         })

split.screen(rbind(c(0, 0.7, 0, 1), c(0.7, 1, 0, 1)))
chain <- c("dotted", "dashed", "blank", "solid")

screen(1); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 2, 2, 1))
for (i in c(1, 2, 4)){
  if (i != 1){ par(new = TRUE) }
  plot(stan_out_tmp[, i], type = "l", lty = chain[i], ann = FALSE, axes = FALSE,
       xlim = c(1, sample_len), ylim = c(sample_min, sample_max))
}
axis(side = 1, xlim = c(1, sample_len))
mtext(text = "Number of searching steps", side = 1, line = 2)
mtext(text = "Sample value", side = 2, line = 0)
legend("topright", legend = c("1st", "2nd", "3rd"), lty = chain[c(1, 2, 4)], cex = 0.6)

screen(2); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 0, 2, 0))
plot(0, type = "n", ann = FALSE, axes = FALSE,
     xlim = c(min(y_val), max(y_val)*2.0), ylim = c(sample_min, sample_max*1.1))
for (i in c(1, 2, 4)){
  polygon(y = c(x_val, rev(x_val)), x = c(y_val[, i], rep(y_val[dens_len, i], dens_len)),
          border = NA, col = rgb(0.5, 0.5, 0.5, 0.5))
  lines(x = y_val[, i], y = x_val, lty = chain[i])
}
text(x = max(y_val), y = median(x_val), adj = -0.2, labels = "Distribution of interest", srt = -90)

close.screen(all = TRUE)
```


##State estimation with MCMC

##Use of library

###Various libraries

###Example: artificial local level model

```{r Code 10.2, collapse=TRUE}
# <<Smoothing for local level model using MCMC (known parameters)>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code, parallel computation, and aspect ratio of plot
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_get() + theme(aspect.ratio = 3/4))

# Load data on artificial local level model
load(file = "ArtifitialLocalLevelModel.RData")

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-1.stan")

# Smoothing: Execution (sampling)
fit_stan <- sampling(object = stan_mod_out,
                     data = list(t_max = t_max, y = y, 
                                 W = mod$W, V = mod$V, 
                                 m0 = mod$m0, C0 = mod$C0),
                     pars = c("x"),
                     seed = 123
            )

# Confirmation of the results
oldpar <- par(no.readonly = TRUE); options(max.print = 99999)
fit_stan
par(oldpar)
traceplot(fit_stan, pars = c(sprintf("x[%d]", 100), "lp__"), alpha = 0.5)

# Extract necessary sampling result
stan_mcmc_out <- rstan::extract(fit_stan, pars = "x")
str(stan_mcmc_out)

# Calculate the mean, 25%, and 75% values while marginalization
s_mcmc <- colMeans(stan_mcmc_out$x)
s_mcmc_quant <- apply(stan_mcmc_out$x, 2, FUN = quantile, probs=c(0.25, 0.75))

# Omit display of the following codes

# Plot results
ts.plot(cbind(y, s), col = c("lightgray", "blue"))
lines(s_mcmc, col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "Mean (Kalman smoothing)",  "Mean (MCMC)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 50, cex = 0.6)

# Plot results
ts.plot(cbind(y, do.call("cbind", s_quant)),
        col = c("lightgray", "blue", "blue"))
lines(s_mcmc_quant["25%", ], col = "red", lty = "dashed")
lines(s_mcmc_quant["75%", ], col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "50% intervals (Kalman smoothing)",  "50% intervals (MCMC)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 60, cex = 0.6)
```



##Estimation example in general state space model

```{r Code 10.4, collapse=TRUE}
# <<Smoothing for local level model using MCMC (unknown parameter)>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code, parallel computation, and aspect ratio of plot
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_get() + theme(aspect.ratio = 3/4))

# Load data on artificial local level model
load(file = "ArtifitialLocalLevelModel.RData")

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-2.stan")

# Smoothing: Execution (sampling)
fit_stan <- sampling(object = stan_mod_out,
                     data = list(t_max = t_max, y = y, 
                                 m0 = mod$m0, C0 = mod$C0),
                     pars = c("W", "V", "x"),
                     seed = 123
            )

# Confirmation of the results
oldpar <- par(no.readonly = TRUE); options(max.print = 99999)
fit_stan

# Omit display of the following codes

par(oldpar)
traceplot(fit_stan, pars = c("W", "V"), alpha = 0.5)

# Extract necessary sampling result
stan_mcmc_out <- rstan::extract(fit_stan, pars = "x")

# Calculate the mean, 25%, and 75% values while marginalization
s_mcmc <- colMeans(stan_mcmc_out$x)
s_mcmc_quant <- apply(stan_mcmc_out$x, 2, FUN = quantile, probs=c(0.25, 0.75))

# Plot results
ts.plot(cbind(y, s), col = c("lightgray", "blue"))
lines(s_mcmc, col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "Mean (Kalman smoothing)",  "Mean (MCMC)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 50, cex = 0.6)

# Plot results
ts.plot(cbind(y, do.call("cbind", s_quant)),
        col = c("lightgray", "blue", "blue"))
lines(s_mcmc_quant["25%", ], col = "red", lty = "dashed")
lines(s_mcmc_quant["75%", ], col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "50% intervals (Kalman smoothing)",  "50% intervals (MCMC)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 60, cex = 0.6)
```


##Technique for improving estimation accuracy

###The case where the linear Gaussian state space model is partially fitted

###Example: artificial local level model

```{r Code 10.6, collapse=TRUE}
# <<Smoothing for local level model using MCMC (unknown parameters and the use of Kalman filter)>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code, parallel computation, and aspect ratio of plot
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_get() + theme(aspect.ratio = 3/4))

# Load data on artificial local level model
load(file = "ArtifitialLocalLevelModel.RData")

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-3.stan")

# Smoothing: Execution (sampling)
dim(mod$m0) <- 1               # Set the explicit dimension in case of only one element vector
fit_stan <- sampling(object = stan_mod_out,
                     data = list(t_max = t_max, y = matrix(y, nrow = 1), 
                                 G = mod$G, F = t(mod$F),
                                 m0 = mod$m0, C0 = mod$C0),
                     pars = c("W", "V"),
                     seed = 123
            )

# Confirmation of the results
fit_stan
traceplot(fit_stan, pars = c("W", "V"), alpha = 0.5)
```




```{r Code 10.7, collapse=TRUE}
# <<Smoothing for local level model using MCMC (state draw with FFBS)>>

# Preprocessing
set.seed(123)
library(dlm)

# Extract necessary sampling result
stan_mcmc_out <- rstan::extract(fit_stan, pars = c("W", "V"))

# FFBS preprocessing: setting of MCMC iteration step and progress bar
it_seq <- seq_along(stan_mcmc_out$V[, 1, 1])
progress_bar <- txtProgressBar(min = 1, max = max(it_seq), style = 3)

# FFBS main processing: draw of state
x_FFBS <- sapply(it_seq, function(it){
  # Display progress bar
  setTxtProgressBar(pb = progress_bar, value = it)

  # Set values of W and V to the model
  mod$W[1, 1] <- stan_mcmc_out$W[it, 1, 1]
  mod$V[1, 1] <- stan_mcmc_out$V[it, 1, 1]

  # FFBS execution
  return(dlmBSample(dlmFilter(y = y, mod = mod)))
})

# FFBS post-processing: removal of x0 and transposition (row means time direction according to the output of Stan)
x_FFBS <- t(x_FFBS[-1, ])

# Calculate the mean, 25%, and 75% values while marginalization
s_FFBS <- colMeans(x_FFBS)
s_FFBS_quant <- apply(x_FFBS, 2, FUN = quantile, probs=c(0.25, 0.75))

# Omit display of the following codes

# Plot results
ts.plot(cbind(y, s), col = c("lightgray", "blue"))
lines(s_FFBS, col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "Mean (Kalman smoothing)",  "Mean (FFBS)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 50, cex = 0.6)

# Plot results
ts.plot(cbind(y, do.call("cbind", s_quant)),
        col = c("lightgray", "blue", "blue"))
lines(s_FFBS_quant["25%", ], col = "red", lty = "dashed")
lines(s_FFBS_quant["75%", ], col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "50% intervals (Kalman smoothing)",  "50% intervals (FFBS)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 60, cex = 0.6)
```




###Example: monthly totals of car drivers in the UK killed or injured

```{r Code 10.8, collapse=TRUE}
# <<Monthly totals of car drivers in the UK killed or injured: Kalman smoothing>>

# Preprocessing
set.seed(123)
library(dlm)

# Log-transform the data, set the time series length
y <- log(UKDriverDeaths)
t_max <- length(y)

# Plot with horizontal axis as time
plot(y)

# Model template
mod <- dlmModPoly(order = 1) + dlmModSeas(frequency = 12)

# User-defined function to define and build a model
build_dlm_UKD <- function(par) {
  mod$W[1, 1] <- exp(par[1])
  mod$W[2, 2] <- exp(par[2])
  mod$V[1, 1] <- exp(par[3])

  return(mod)
}

# Maximum likelihood estimation of parameters
fit_dlm_UKD <- dlmMLE(y = y, parm = rep(0, times = 3), build = build_dlm_UKD)

# Model setting and its confirmation
mod <- build_dlm_UKD(fit_dlm_UKD$par)
cat(diag(mod$W)[1:2], mod$V, "\n")

# Smoothing processing
dlmSmoothed_obj <- dlmSmooth(y = y, mod = mod)

# Mean of the smoothing distribution
   mu <- dropFirst(dlmSmoothed_obj$s[, 1])
gamma <- dropFirst(dlmSmoothed_obj$s[, 2])

# Plot results
oldpar <- par(no.readonly = TRUE)
par(mfrow = c(3, 1)); par(oma = c(2, 0, 0, 0)); par(mar = c(2, 4, 1, 1))
ts.plot(    y, ylab = "Observations (log-transformed)")
ts.plot(   mu, ylab = "Level component")
ts.plot(gamma, ylab = "Seasonal component")
mtext(text = "Time", side = 1, line = 1, outer = TRUE)
par(oldpar)
```




```{r Code 10.10, collapse=TRUE}
# <<Monthly totals of car drivers in the UK killed or injured: state sampling with MCMC>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code, parallel computation, and aspect ratio of plot
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_get() + theme(aspect.ratio = 3/4))

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-4.stan")

# Smoothing: Execution (sampling)
fit_stan <- sampling(object = stan_mod_out,
                     data = list(n = t_max, y = y, m0 = mod$m0, C0 = mod$C0), 
                     pars = c("W_mu", "W_gamma", "V"),
                     seed = 123
            )

# Confirmation of the results
fit_stan
traceplot(fit_stan, pars = c("W_mu", "W_gamma", "V"), alpha = 0.5)
```




```{r Code 10.12, collapse=TRUE}
# <<Monthly totals of car drivers in the UK killed or injured: no state sampling with MCMC>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code, parallel computation, and aspect ratio of plot
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
theme_set(theme_get() + theme(aspect.ratio = 3/4))

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-5.stan")

# Smoothing: Execution (sampling)
fit_stan <- sampling(object = stan_mod_out,
                     data = list(t_max = t_max, y = matrix(y, nrow = 1), 
                                 G = mod$G, F = t(mod$F),
                                 m0 = mod$m0, C0 = mod$C0),
                     pars = c("W_mu", "W_gamma", "V"),
                     seed = 123
            )

# Confirmation of the results
fit_stan
traceplot(fit_stan, pars = c("W_mu", "W_gamma", "V"), alpha = 0.5)

# Confirm the mean of marginal distribution as estimation result
cat(summary(fit_stan)$summary[   "W_mu", "mean"],
    summary(fit_stan)$summary["W_gamma", "mean"],
    summary(fit_stan)$summary[ "V[1,1]", "mean"], "\n")
```




```{r Code 10.13, collapse=TRUE}
# <<Monthly totals of car drivers in the UK killed or injured: state draw with FFBS>>

# Preprocessing
set.seed(123)
library(dlm)

# Extract necessary sampling result
stan_mcmc_out <- rstan::extract(fit_stan, pars = c("W_mu", "W_gamma", "V"))

# FFBS preprocessing: setting of MCMC iteration step and progress bar
it_seq <- seq_along(stan_mcmc_out$V[, 1, 1])
progress_bar <- txtProgressBar(min = 1, max = max(it_seq), style = 3)

# FFBS main processing: draw of state
x_FFBS <- lapply(it_seq, function(it){
  # Display progress bar
  setTxtProgressBar(pb = progress_bar, value = it)

  # Set values of W and V to the model
  mod$W[1, 1] <- stan_mcmc_out$W_mu[it]
  mod$W[2, 2] <- stan_mcmc_out$W_gamma[it]
  mod$V[1, 1] <- stan_mcmc_out$V[it, 1, 1]

  # FFBS execution
  return(dlmBSample(dlmFilter(y = y, mod = mod)))
})

# FFBS post-processing: removal of x0 and transposition (row means time direction according to the output of Stan)
   x_mu_FFBS <- t(sapply(x_FFBS, function(x){ x[-1, 1] }))
x_gamma_FFBS <- t(sapply(x_FFBS, function(x){ x[-1, 2] }))

# Calculate the mean while marginalization
   mu_FFBS <- colMeans(   x_mu_FFBS)
gamma_FFBS <- colMeans(x_gamma_FFBS)

# Omit display of the following codes

# Cast to ts class
   mu_FFBS <- ts(   mu_FFBS); tsp(   mu_FFBS) <- tsp(   mu)
gamma_FFBS <- ts(gamma_FFBS); tsp(gamma_FFBS) <- tsp(gamma)

# Plot results
ts.plot(cbind(y, mu), col = c("lightgray", "blue"))
lines(mu_FFBS, col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "Mean (Kalman smoothing)",  "Mean (FFBS)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright",  cex = 0.6)

# Plot results
ts.plot(gamma, col = "blue", ylab = "", ylim = c(-0.15, 0.31))
lines(gamma_FFBS, col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "Mean (Kalman smoothing)",  "Mean (FFBS)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", cex = 0.6)
```










```{r Post-processing for pdf plot, echo = FALSE, include = FALSE}
# <<Post-processing for pdf plot>>

if (SAVE_PLOT_PDF == TRUE){
  showtext.end()

  dev.off()
}
```
