---
title: "Code"
output:
  html_document:
    number_section: true
    pandoc_args: [--number-offset=3]
---


```{r For setup, echo = FALSE, include = FALSE}
# Working directory ... set your favorite
setwd("C:/cygwin/home/NOC/ebook/edoc")

# Character width in console output
options(width = 84)

# Plot related setting
SAVE_PLOT_PDF <- F
if (SAVE_PLOT_PDF == TRUE){
  # Exporting to PDF file
  pdf(height = 7 / (4/3))

  # Rasterize the font information
  require(showtext)
  font.add("meiryo", "meiryo.ttc")
  par(family = "meiryo")
  showtext.begin()
}
```





#Quick tour of time series analysis

##Confirmation of the purpose and data collection

##Preliminary examination of data

###Plot with horizontal axis as time

```{r Code 4.1, collapse=TRUE}
# <<Plot with horizontal axis as time>>

# Preprocessing about plot (saving the default settings for plot, then changing them)
oldpar <- par(no.readonly = TRUE)
par(mfrow = c(2, 2)); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 4, 2, 1))

# (a) Annual flow of the Nile
plot(Nile, main = "(a)")

# (b) CO2 concentration in the atmosphere
# Load the data
Ryori <- read.csv("CO2.csv")

# Cast the data to ts class, truncating it by December 2014
y_all <- ts(data = Ryori$CO2, start = c(1987, 1), frequency = 12)
y <- window(y_all, end = c(2014, 12))
y_CO2 <- y
plot(y_CO2, main = "(b)")

# (c) Quarterly gas consumption in the UK
plot(UKgas, main = "(c)")

# (d) Artificial generated data from a nonlinear model
load("BenchmarkNonLinearModel.RData")
y_nonlinear <- ts(y)
plot(y_nonlinear, main = "(d)")

# Post-processing about plot
par(oldpar)
```



```{r Figure 4.2, echo = FALSE}
# <<Plot with horizontal axis as time (log-transformed data)>>


# (c) Quarterly gas consumption in the UK (log-transformed data)
UKgas_log <- log(UKgas)
plot(UKgas_log, main = "Log of (c)")
```


###Histogram and five number summary

```{r Code 4.2, collapse=TRUE}
# <<Histogram and five number summary>>

# Preprocessing about plot (saving the default settings for plot, then changing them)
oldpar <- par(no.readonly = TRUE)
par(mfrow = c(2, 2)); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 4, 2, 1))

# (a) Annual flow of the Nile
hist(Nile, main = "(a)", xlab = "Data Value")
summary(Nile)

# (b) CO2 concentration in the atmosphere
hist(y_CO2, main = "(b)", xlab = "Data Value")
summary(y_CO2)

# (c) Quarterly gas consumption in the UK
hist(UKgas_log, main = "(c)", xlab = "Data Value")
summary(UKgas_log)

# (d) Artificial generated data from a nonlinear model
hist(y_nonlinear, main = "(d)", xlab = "Data Value")
summary(y_nonlinear)

# Post-processing about plot
par(oldpar)
```



```{r Figure 4.4, echo = FALSE, results='hide'}
# <<Outlier>>

# Preprocessing about plot (saving the default settings for plot, then changing them)
oldpar <- par(no.readonly = TRUE)
par(mfrow = c(1, 2), pty = "s"); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 4, 2, 1))

# Tokyo's average temperature in August of particular year
temperature_without_outl <- c(25.6, 25.6, 24.9, 25.1, 24.9, 25.7, 27.2, 28, 29.1, 30.2, 30.5, 30.6, 30.6, 30.2, 30.4, 29.6, 28.6, 27.9, 27.3, 27.3, 26.9, 26.5, 26.5)
temperature_with_outl    <- c(25.6, 25.6, 24.9, 25.1, 24.9, 25.7, 27.2, 28, 29.1, 30.2, 30.5, 30.6, 30.6, 30.2, 30.4, 29.6, 28.6, 27.9, 27.3, 67.3, 26.9, 26.5, 26.5)

# Plot
ts.plot(temperature_with_outl, ylim = c(20, 40), type = "l", ylab = "Celsius temperature")

# Histogram and five number summary
hist(temperature_with_outl, breaks = seq(from = 20, to = 80, by = 1), 
     xlab = "Celsius temperature", ylab = "Frequency", main = "")
summary(temperature_with_outl)

# Post-processing about plot
par(oldpar)
```



```{r Code 4.3, collapse=TRUE}
# <<Compensation for NA>>

# (b) CO2 concentration in the atmosphere

# Identification of NA location
NA.point <- which(is.na(y_CO2))

# Compensation for NA: arithmetic mean about before and after the NA
y_CO2[NA.point] <- (y_CO2[NA.point-1] + y_CO2[NA.point+1]) / 2
```


###Autocorrelation coefficients


```{r Code 4.4, collapse=TRUE}
# <<Autocorrelation coefficient>>

# Preprocessing about plot (saving the default settings for plot, then changing them)
oldpar <- par(no.readonly = TRUE)
par(mfrow = c(2, 2)); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 4, 3, 1))


# (a) Annual flow of the Nile
acf(Nile, main = "(a)")

# (b) CO2 concentration in the atmosphere
acf(y_CO2, main = "(b)")

# (c) Quarterly gas consumption in the UK
acf(UKgas_log, main = "(c)")

# (d) Artificial generated data from a nonlinear model
acf(y_nonlinear, main = "(d)")

# Post-processing about plot
par(oldpar)
```


###Frequency spectrum

####Fundamentals on frequency


```{r Figure 4.6, echo = FALSE, results='hide'}
# <<Fourier series expansion>>

# Preprocessing about plot (saving the default settings for plot, then changing them)
oldpar <- par(no.readonly = TRUE)
split.screen(rbind(c(0, 0.6, 0.75, 1   ),
                   c(0, 0.6, 0.50, 0.75),
                   c(0, 0.6, 0.25, 0.50),
                   c(0, 0.6, 0   , 0.25)))

# Tick values on the x axis
x_tick <- seq(from = 0, to = 2*pi, length.out = 1000)

# Function drawing rectangular signal
draw_rectangle <- function(){
  lines(x = c(   0,    0), y = c( 0,  1))
  lines(x = c(   0, 1*pi), y = c( 1,  1))
  lines(x = c(1*pi, 1*pi), y = c( 1, -1))
  lines(x = c(1*pi, 2*pi), y = c(-1, -1))
  lines(x = c(2*pi, 2*pi), y = c(-1,  0))
}

# Fourier series expansion
Fourie_series <- 0
for (n in c(1, 3, 5, 7)){
  screen((n+1)/2); par(oma = c(0, 0, 0, 0)); par(mar = c(1, 0, 0, 0))
  
  Fourie_series <- Fourie_series + 4/pi * sin(n * x_tick) / n
  plot(x = x_tick, y = Fourie_series, ylim = c(-1.3, 1.3), 
       type = "l", lty = "dashed", ann = FALSE, xaxt="n")
  draw_rectangle()
  axis(side = 1, at = 0:6, labels = FALSE, tcl = -0.2)
  mtext(text = 0:6, at = 0:6, side = 1, line = 0, cex = 0.7)
  legend("topright", legend = sprintf("Sum of %d sin()s", ceiling(n/2)), cex = 0.6)
}

# Post-processing about plot
close.screen(all = TRUE)
par(oldpar)
```


####Frequency domain transform


```{r Code 4.5, collapse=TRUE}
# <<Frequency domain transform>>

# User-defined function drawing frequency spectrum (tick: marking points on horizontal axis, unit: quantity between time points)
plot.spectrum <- function(dat, lab = "", main = "",
                          y_max = 1, tick = c(8, 4), unit = 1){
  # Frequency domain transform of data
  dat_FFT <- abs(fft(as.vector(dat)))

  # Preparation for display setting about horizontal axis (frequency)
  data_len  <- length(dat_FFT)
  freq_tick <- c(data_len, tick, 2)

  # Plot data in frequency domain
  plot(dat_FFT/max(dat_FFT), type = "l", main = main,
       ylab = "|Standardized frequency spectrum|", ylim = c(0, y_max),
       xlab = sprintf("Frequency [1/%s]", lab), xlim = c(1, data_len/2), xaxt = "n")
  axis(side = 1, at = data_len/freq_tick * unit + 1, 
       labels = sprintf("1/%d", freq_tick), cex.axis = 0.7)
}


# Preprocessing about plot (saving the default settings for plot, then changing them)
oldpar <- par(no.readonly = TRUE)
par(mfrow = c(2, 2)); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 4, 3, 1))

# (a) Annual flow of the Nile
plot.spectrum(       Nile, lab =   "Year", main = "(a)")

# (b) CO2 concentration in the atmosphere
plot.spectrum(      y_CO2, lab =   "Month", main = "(b)", tick = c(12, 6))

# (c) Quarterly gas consumption in the UK
plot.spectrum(  UKgas_log, lab =   "Month", main = "(c)", tick = c(12, 6), unit = 3)

# (d) Artificial generated data from a nonlinear model
plot.spectrum(y_nonlinear, lab = "Time point", main = "(d)")

# Omit display of the following codes

# Redraw with changing the scale of vertical axis

# (a) Annual flow of the Nile
plot.spectrum(       Nile, lab =   "Year", main = "(a)", y_max = 0.02)

# (b) CO2 concentration in the atmosphere
plot.spectrum(      y_CO2, lab =   "Month", main = "(b)", y_max = 0.02,
              tick = c(12, 6))

# (c) Quarterly gas consumption in the UK
plot.spectrum(  UKgas_log, lab =   "Month", main = "(c)", y_max = 0.2 ,
              tick = c(12, 6), unit = 3)

# (d) Artificial generated data from a nonlinear model
plot.spectrum(y_nonlinear, lab = "Time point", main = "(d)", y_max = 0.2 )

# Post-processing about plot
par(oldpar)
```



##Model definition

```{r Figure 4.10, echo = FALSE, results='hide'}
# <<Decomposition and residuals of time series data>>

# Preprocessing
set.seed(1234)
oldpar <- par(no.readonly = TRUE)
par(cex.lab = 1.0)
par(mfrow = c(4, 3)); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 4, 2, 1))

# Tick on the horizontal axis and noise
x_tick <- seq(from = 0, to = 2*pi*5, length = 100)
noise <- rnorm(n = 100) / 3

# Level component
plot.new()  # Dummy panel
plot(x = x_tick, y = rep(10, 100),
     type = "l", bty = "l", xaxs = "r", yaxs = "i", xaxt = "n", yaxt = "n", ann = F)
mtext(text = "Level component", side = 1, line = 1.5)
mtext(text = "Time", side = 1, at = 10*pi, line = 0.3, cex = 0.7)
plot.new()  # Dummy panel

# Trend component
plot.new()  # Dummy panel
plot(x = x_tick, y = x_tick/10, ylim = c(0, 10),
     type = "l", bty = "l", xaxs = "i", yaxs = "i", xaxt = "n", yaxt = "n", ann = F)
mtext(text = "Trend component", side = 1, line = 1.5)
mtext(text = "Time", side = 1, at = 10*pi, line = 0.3, cex = 0.7)
plot.new()  # Dummy panel

# Seasonal component
plot.new()  # Dummy panel
plot(x = x_tick, y = sin(x_tick),
     type = "l", bty = "l", xaxs = "r", yaxs = "r", xaxt = "n", yaxt = "n", ann = F)
mtext(text = "Seasonal component", side = 1, line = 1.5)
mtext(text = "Time", side = 1, at = 10*pi, line = 0.3, cex = 0.7)
plot.new()  # Dummy panel

# Time series data
plot(x = x_tick, y = 10 + x_tick/10 + sin(x_tick) + noise, ylim = c(5, 14),
     type = "l", bty = "l", xaxs = "r", yaxs = "i", xaxt = "n", yaxt = "n", ann = F)
mtext(text = "Time series data", side = 1, line = 1.5)
mtext(text = "Time", side = 1, at = 10*pi, line = 0.3, cex = 0.7)

# Estimates
plot(x = x_tick, y = 10 + x_tick/10 + sin(x_tick), ylim = c(5, 14),
     type = "l", bty = "l", xaxs = "r", yaxs = "i", xaxt = "n", yaxt = "n", ann = F)
mtext(text = "Estimates", side = 1, line = 1.5)
mtext(text = "Time", side = 1, at = 10*pi, line = 0.3, cex = 0.7)

# Residuals
plot(x = x_tick, y = noise, ylim = c(-3, 3),
     type = "l", bty = "n", xaxs = "i", yaxs = "i", xaxt = "n", yaxt = "n", ann = F)
abline(v = 0, h = 0)
mtext(text = "Residuals", side = 1, line = 1.5)
mtext(text = "Time", side = 1, at = 10*pi, line = -1, cex = 0.7)

# Post-processing
par(oldpar)
```


##Specification of parameter values

##Execution of filtering, prediction, and smoothing



```{r Code 4.6, collapse=TRUE}
# <<Holt--Winters method>>

# (a) Annual flow of the Nile
HW_Nile <- HoltWinters(Nile, beta = FALSE, gamma = FALSE)
str(HW_Nile)

# (b) CO2 concentration in the atmosphere
HW_CO2 <- HoltWinters(y_CO2)

# (c) Quarterly gas consumption in the UK
HW_UKgas_log <- HoltWinters(UKgas_log)

# (d) Artificial generated data from a nonlinear model
HW_nonlinear <- HoltWinters(y_nonlinear, gamma = FALSE)

# Plot the filtering value

# Preprocessing about plot (saving the default settings for plot, then changing them)
oldpar <- par(no.readonly = TRUE)
par(mfrow = c(2, 2)); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 4, 3, 1))
mygray <- "#80808080"

plot(HW_Nile     , main = "(a)",
     col = mygray, col.predicted = "black", lty.predicted = "dashed")
plot(HW_CO2      , main = "(b)",
     col = mygray, col.predicted = "black", lty.predicted = "dashed")
plot(HW_UKgas_log, main = "(c)",
     col = mygray, col.predicted = "black", lty.predicted = "dashed")
plot(HW_nonlinear, main = "(d)",
     col = mygray, col.predicted = "black", lty.predicted = "dashed")

# Post-processing about plot
par(oldpar)
```


```{r Code 4.7, collapse=TRUE}
# <<Holt--Winters method (per component)>>

# (a) Annual flow of the Nile
HW_out <- HW_Nile
HW_decomp <- ts.union(y = HW_out$x,
                  Level = HW_out$fitted[,  "level"], 
              Resisuals = residuals(HW_out))
plot(HW_decomp, main = "")

# (b) CO2 concentration in the atmosphere
HW_out <- HW_CO2
HW_decomp <- ts.union(y = HW_out$x,
                  Level = HW_out$fitted[,  "level"] + HW_out$fitted[, "trend"], 
                 Season = HW_out$fitted[, "season"],
              Resisuals = residuals(HW_out))
plot(HW_decomp, main = "")

# (c) Quarterly gas consumption in the UK
HW_out <- HW_UKgas_log
HW_decomp <- ts.union(y = HW_out$x,
                  Level = HW_out$fitted[,  "level"] + HW_out$fitted[, "trend"], 
                 Season = HW_out$fitted[, "season"],
              Resisuals = residuals(HW_out))
plot(HW_decomp, main = "")

# (d) Artificial generated data from a nonlinear model
HW_out <- HW_nonlinear
HW_decomp <- ts.union(y = HW_out$x,
                  Level = HW_out$fitted[,  "level"] + HW_out$fitted[, "trend"], 
              Resisuals = residuals(HW_out))
plot(HW_decomp, main = "")
```


```{r Code 4.8, collapse=TRUE}
# <<Holt--Winters method (forecast)>>

# (b) CO2 concentration in the atmosphere
HW_predict <- predict(HW_CO2, n.ahead = 12)
str(HW_predict)

# Plot observations along with filtering and prediction values
plot(HW_CO2, HW_predict, main = "Filtering and prediction with Holt--Winters method",
     col = mygray, col.predicted = "black", lty.predicted = "dashed")

# Plot also observations in 2015
y_CO2_2015 <- window(y_all, start = 2015)
lines(y_CO2_2015, col = mygray)
```



##Diagnostic checking for the results

```{r Code 4.9, collapse=TRUE}
# <<Diagnostic checking for the results (autocorrelation of residuals)>>

# Preprocessing about plot (saving the default settings for plot, then changing them)
oldpar <- par(no.readonly = TRUE)
par(mfrow = c(2, 2)); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 4, 3, 1))

acf(residuals(HW_Nile)     , main = "(a)")
acf(residuals(HW_CO2)      , main = "(b)")
acf(residuals(HW_UKgas_log), main = "(c)")
acf(residuals(HW_nonlinear), main = "(d)")

# Post-processing about plot
par(oldpar)
```


```{r Code 4.10, collapse=TRUE}
# <<Diagnostic checking for the results (prediction performance)>>

# User-defined function for mean absolute percentage error (MAPE)
MAPE <- function(true, pred){
  mean(abs(pred - true) / true)
}

# The mean absolute percentage error of the prediction value
MAPE(true = y_CO2_2015, pred = HW_predict)
```










```{r Post-processing for pdf plot, echo = FALSE, include = FALSE}
# <<Post-processing for pdf plot>>

if (SAVE_PLOT_PDF == TRUE){
  showtext.end()

  dev.off()
}
```
